---
title: "Cam Report"
author: "Lyron Winderbaum"
date: "9 August 2017"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(reshape2)
knitr::opts_chunk$set(echo = TRUE)
```

## The Data

Here I'll go through a little dataset I collected in which the subjects are 
cams. Below is an image of a cam in use -- a size 6 Black Diamond Camalot C4, to 
be precise. A cam is a device mostly used in recreational rockclimbing that can 
be manipulated to have a variety of different widths and hence be placed into 
cracks of different sizes.

![A #6 Black Diamond Camalot C4 placed in a crack (image taken from [wikipedia](https://en.wikipedia.org/wiki/Camalot)).](https://upload.wikimedia.org/wikipedia/commons/thumb/f/f7/Camalot_number_6.JPG/1200px-Camalot_number_6.JPG)

There are several differernt brands that manufacture cams, and some of them 
provide useful charts such as the one below.

![A cam comparison chart taken from [the black diamond website](http://blackdiamondequipment.com/en/climbing-cams-stoppers-nuts-hexes/camalottm-BD262104_cfg.html)).](https://blackdiamondequipment.com/on/demandware.static/-/Library-Sites-SharedLibrary/default/dwaf34dfb1/images/tech-highlights/tech-highlight_S16_cam-chart.png)

Now this chart is great, but not all brands provide such a nice figure, 
and nobody provides a nice figure comparing different brands, or 
organising it differently to put similar sized cams next to each other, or 
looking at other factors like weight, etc. So I decided to go around to each of 
the brands websites and collect all the technical specifications and 
curate them into as single dataset, so that I could:

 - Figure out how to make weird plots like this, which show a range, using 
 ggplot2 --- something I thought was pretty interesting.
 - Make other, more useful plots comparing different brands and placing 
 cams of similar sizes next to each other so they can more easily be compared.

So I did that, collected technical specifcations for 196 cams from four
different brands:

```{r read_data}
cam_df = read.csv('cam_df.csv', stringsAsFactors = FALSE)
head(cam_df)
```

Each row of this data.frame represents a different type or size of cam, and here 
I will give a more detailed description of each of the columns, or variables, 
recorded for each cam:

 - **Brand** --- The company that manufactured the cam. I collected 
 technical specifications from four brands: BD (Black Diamond), 
 Totemcam, Metolius, and DMM.
 - **Variety** --- a single brand often produces different varieties of 
 cam, and this indicates this subdivision.
 - **NoCams** --- indicates the number of camming units in the cam, usually either
 three or four. 
 - **Size** --- Usually a number, with the exception of the Metolius Super cams, 
 used to identify different sizes within a single variety of cam, typically 
 larger numbers indicate bigger cams.
 - **Colour** --- Often the different sizes are colour coded, although different 
 brands can have different colour schemes.
 - **Weight** --- Weight in grams (g).
 - **Strenth** --- Rated strength in kilo-Newtons (kN).
 - **Min** --- Minimum width in centimeters (cm).
 - **Max** --- Maximum width in centimeters (cm).

Note that the values in the **Colour** column correspond to the colours as 
reported in the technical specifcations, which when given to ggplot2 as strings
are interpreted as colours, but not neccessarily colours that actually look 
much like the colours of the cams themselves, so I'll make a few replacements
to better match the look of the actual colours.

```{r adjust_colours}
# Fluro green and lighter grey look bad
cam_df[cam_df$Colour == "green",        "Colour"] = "darkgreen"
cam_df[cam_df$Colour == "grey",         "Colour"] = "darkgrey"

# Metolius `light blue', `dark blue' and `light purple'
cam_df[cam_df$Colour == "light blue",   "Colour"] = "skyblue"
cam_df[cam_df$Colour == "light purple", "Colour"] = "orchid"
cam_df[cam_df$Colour == "dark blue",    "Colour"] = "darkblue"

# DMM `silver' (very dark grey)
cam_df[cam_df$Colour == "silver",       "Colour"] = "grey50"
```

For convenience I'll also create a unique identifier column by combining the 
**Brand**, **Variety** and **Size** columns.

```{r unique_id}
cam_df$ID = paste(cam_df$Brand, cam_df$Variety, cam_df$Size)
```

## Plots

```{r plot_dragon_vs_camalot}
# Compare C4s with corresponding sized Dragons
cam_subset = subset(cam_df, Variety %in% c('C4', 'Dragon'))
# Remove the two biggest C4 cams as they are bigger than the biggest Dragon Cams
cam_subset = subset(cam_subset, !((Variety == 'C4') & (Size %in% c('5', '6'))))
# Sort factor values neatly.
cam_subset = cam_subset[order(cam_subset$Brand, cam_subset$Min), ]

# using annotate()
n = nrow(cam_subset)
p <- ggplot()
p <- p + annotate("rect", xmin = (n - 0.5):0.5, xmax = (n + 0.5):1.5,
                  ymin = cam_subset$Min, ymax = cam_subset$Max,
                  alpha = 1, fill = cam_subset$Colour)
p <- p + annotate("text", x = n:1, y = (cam_subset$Min + cam_subset$Max) / 2,
                  label = cam_subset$ID, size = 2)
p <- p + scale_x_continuous(name = '', breaks = 1:n, 
                            labels = rev(cam_subset$ID))
p <- p + scale_y_continuous(name = "Range (mm)")
p <- p + theme(panel.grid.minor = element_blank())
p <- p + coord_flip()
print(p)
```

Nice. This is a comparison of DMM Dragon cams to BD Camalots, in the sizes in 
which they have comparable cams. But now that we have it in a data frame we 
can reorganise this to make a better plot:

```{r plot_dragon_vs_camalot_resorted}
# Sort factor values better.
tmp.order = ceiling((1:n)/2)
tmp.order[seq(2, n, 2)] = ((n / 2) + 1):n
cam_subset = cam_subset[tmp.order, ]

# using annotate()
n = nrow(cam_subset)
p <- ggplot()
p <- p + annotate("rect", xmin = (n - 0.5):0.5, xmax = (n + 0.5):1.5,
                  ymin = cam_subset$Min, ymax = cam_subset$Max,
                  alpha = 1, fill = cam_subset$Colour)
p <- p + annotate("text", x = n:1, y = (cam_subset$Min + cam_subset$Max) / 2,
                  label = cam_subset$ID, size = 2)
p <- p + scale_x_continuous(name = '', breaks = 1:n, 
                            labels = rev(cam_subset$ID))
p <- p + scale_y_continuous(name = "Range (mm)")
p <- p + theme(panel.grid.minor = element_blank())
p <- p + coord_flip()
print(p)
```

and we could plot weight on the y-axis, instead of ID:

```{r plot_dragon_vs_camalot_weight}
# using annotate()
n = nrow(cam_subset)
p <- ggplot()
p <- p + annotate("rect", xmin = cam_subset$Weight - 4, 
                  xmax = cam_subset$Weight + 4,
                  ymin = cam_subset$Min, ymax = cam_subset$Max,
                  alpha = 0.4, fill = cam_subset$Colour)
p <- p + annotate("text", x = cam_subset$Weight, 
                  y = (cam_subset$Min + cam_subset$Max) / 2,
                  label = cam_subset$ID, size = 2)
p <- p + scale_x_continuous(name = 'Weight (g)')
p <- p + scale_y_continuous(name = "Range (mm)")
p <- p + coord_flip()
print(p)
```

we could zoom in a little and do a plot of all the cams in certain size range:

```{r plot_weight_small_cams}
cam_subset = subset(cam_df, Max < 20)
# using annotate()
n = nrow(cam_subset)
p <- ggplot()
p <- p + annotate("rect", xmin = cam_subset$Weight - 0.8, 
                  xmax = cam_subset$Weight + 0.8,
                  ymin = cam_subset$Min, ymax = cam_subset$Max,
                  alpha = 0.4, fill = cam_subset$Colour)
p <- p + annotate("text", x = cam_subset$Weight, 
                  y = (cam_subset$Min + cam_subset$Max) / 2,
                  label = cam_subset$ID, size = 2)
p <- p + scale_x_continuous(name = 'Weight (g)')
p <- p + scale_y_continuous(name = "Range (mm)")
p <- p + coord_flip()
print(p)
```


## Colophon

This report was written in [Rmarkdown](http://rmarkdown.rstudio.com/) using
[RStudio](https://www.rstudio.com/products/rstudio/). 
Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word 
documents. [knitr](https://yihui.name/knitr/) and 
[pandoc](http://pandoc.org/) converted the raw Rmarkdown to html. This colophon 
was modelled after that of the book [R packages](http://r-pkgs.had.co.nz/).

```{r colophon}
devtools::session_info()
```


